A thesis draft

### 2018-11-30 res/sections/08-Background.tex
Fix typos and english

\paragraph*{Architectural view}
About the architectural view of SFCs, proposal defines several assumptions on
them, starting from considering every element of the chain an opaque processing
unit. Another important assumption is that there is no global or standard SF
chaining logic~\cite{rfc7665}. SFC can be easily seen as a graph, because some
SF can have multiple address to forward the elaborated data, thus creating what
is defined as \emph{fork} that ends up forming different branches. This could
lead to loops, that must be treated consequently.

@@

\noindent Regarding this field, the standard does not make any assumption about
the content, and it is left free to be filled with custom metadata. The only
requirement is to set this filed with zeroes when is not used.

>>>>

\paragraph*{Architectural view}
About the SFCs architectural view, the proposal defines several assumptions on
them, starting from considering every element of the chain an opaque processing
unit. Another important assumption is that there is no global or standard SF
chaining logic~\cite{rfc7665}. An SFC can be easily seen as a graph, because
some SF can have multiple address to forward the elaborated data, thus creating
what is defined as \emph{fork} that ends up forming different branches. This
could lead to loops, that must be treated consequently.

@@

\noindent Regarding this field, the standard does not make any assumption about
the content, and it is left free to be filled with custom metadata. The only
requirement is to set this field with zeroes when is not used.

### 2018-11-30 res/sections/09-RelatedWork.tex
Fix typos and english

One of the first proposal that take advantage of container in to take on network
edges NFV is~\cite{cziva2017container}. The authors present the work, based on
previous definition of the platform defined in~\cite{cziva2015container} to
orchestrate the virtual function and to deploy it as close as possible to the
data sender to classify traffic and choose the appropriate functions
immediately.

@@

EPC and ETSI MANO definitions are not the only solutions to have been explored
at this time. An SDN/NFV-based end-to-end network slicing for 5G has been
created in~\cite{chartsias2017sdn}, where the authors provided examples to
demonstrate how end-to-end services can be provisioned using SDN networking, in
particular with controllers like such as OpenDaylight, Pox, Ryu ONOD and
Floodlight. The last one was studied by us too during the technologies analysis,
and a description is available in
Section~\ref{chap:prjan:sec:tech:sub:other:sub:floodlight}.

Compared to our solution, these implementations does not present an unified
framework that gives the possibility orchestrate different network slices
consistently. Adopting Kubernetes gives the advantage to use the Kubernetes
Federation feature~\cite{kubeFederation}, enabling the possibility to simplify
the management of multiple network slice in end-to-end services.

@@

Another important aspect related to 5G networking and network slice is the one
regarding security. Security in this context plays a fundamental role, since
secure communications depends on the security of the overall system.
In~\cite{kotulski2017end} the authors carry out a deep analysis about the
network slice security, putting an emphasis on how this new element (the network
slice) brings additional security questions and problems, especially about
isolation. Moreover, slicing in 5G introduces the concept of shared services,
that is contrary with the definition of isolation itself. In particular,
resource isolation gives the possibility to avoid attacks such us slice
resources exhaustion. Additionally, the authors give a critic regarding the ETSI
NFV-MANO standard making too general assumptions regarding the whole system.
Finally, a Service-Level Agreement (SLA) based orchestration criteria between
network slices is proposed as a viable solution to solve the security problems
about isolation. We tried to mitigate the security concern presented above using
the integrated Role Base Access Control system (RBAC) and other security policy
bundled in by default in Kubernetes.

\section{Provision capabilities}
We noted how, in some other NFV/ETSI-MANO proposals like~\cite{soares2015toward}
and~\cite{abujoda2015midas}, the implementations relays on a static provision
system, where the SFC paths are somehow defined at deploy time, and can't change
without decommissioning, editing and deploying the chain again. Our proposal,
instead, is able to support chain editing on-the-fly, allowing NFV to
dynamically change the SFP.

>>>>

One of the first proposals that takes advantage of the container technology and
to take to the network edges the NFV is~\cite{cziva2017container}. The authors
present the work, based on previous definition of the platform defined
in~\cite{cziva2015container} to orchestrate the virtual function and to deploy
it as close as possible to the data sender to classify traffic and choose the
appropriate functions immediately.

@@

EPC and ETSI MANO definitions are not the only solutions to have been explored
at this time. An SDN/NFV-based end-to-end network slicing for 5G has been
created in~\cite{chartsias2017sdn}, where the authors provided examples to
demonstrate how end-to-end services can be provisioned using SDN networking, in
particular with controllers like OpenDaylight, Pox, Ryu ONOD and Floodlight. The
last one was studied by us too during the technologies analysis, and a
description is available in
Section~\ref{chap:prjan:sec:tech:sub:other:sub:floodlight}.

Compared to our solution, these implementations do not present an unified
framework that gives the possibility orchestrate different network slices
consistently. Adopting Kubernetes gives the advantage to use the Kubernetes
Federation feature~\cite{kubeFederation}, enabling the possibility to simplify
the management of multiple network slice in end-to-end services.

@@

Another important aspect related to 5G networking and network slice is the one
regarding security. Security in this context plays a fundamental role, since
secure communications depends on the security of the overall system.
In~\cite{kotulski2017end} the authors carry out a deep analysis about the
network slice security, putting an emphasis on how this new element (the network
slice) brings additional security questions and problems, especially about
isolation. Moreover, slicing in 5G introduces the concept of shared services,
that is contrary with the definition of isolation itself. In particular,
resource isolation gives the possibility to avoid attacks such us slice
resources exhaustion. Additionally, the authors give a critic regarding the ETSI
NFV-MANO standard making too general assumptions regarding the whole system.
Finally, a Service-Level Agreement (SLA) based orchestration criteria between
network slices is proposed as a viable solution to solve the security problems
about isolation. We tried to mitigate the security concern presented above using
the integrated Role Base Access Control system (RBAC) and other security policy
bundled by default in Kubernetes.

\section{Provision capabilities}
We noticed how, in some other NFV/ETSI-MANO proposals
like~\cite{soares2015toward} and~\cite{abujoda2015midas}, the implementations
relay on a static provision system, where the SFC paths are somehow defined at
deploy time, and can not change without decommissioning, editing and deploying
the chain again. Our proposal, instead, is able to support chain editing
on-the-fly, allowing NFV to dynamically change the SFP.

### 2018-11-30 res/sections/10-ProjectAnalysis.tex
Fix typos and english

As in chapter~\ref{chap:intro}, the creation of a full ETSI MANO compliant
architecture, following the suggestions described in RFC 7665, require the
integration of multiple tools (some of them have been already briefly introduced
in~\ref{chap:intro}). The tools were chosen after their analysis, where pro and
cons have been studied and compared to the necessity of this thesis. Since the
project has been developed with the cloud provided by the University of Padua,
Openstack (which a product analysis can be found later
in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.

@@

\subsubsection{Openstack}
\label{chap:prjan:sec:openstack}
Created in 2010 by a joint operation between Rackspace Hosting and
NASA~\cite{openstackWebsite}, Openstack allows on-premise cloud installations,
using bare-metal resources to provide common cloud services as object storage,
virtual machine deploy, virtual networking. Furthermore its modularity offers
the possibility to add additional components, even proprietary, to achieve a
complete cloud solution.
\begin{figure}[h]
 \centering \includegraphics[scale=0.58]{openstack_logo}
 \caption{Openstack logo}
 \label{chap:prjan:img:openstack_logo}
\end{figure}

@@

As already introduced before, Kubernetes is an open source software framework
specialized in container management and orchestration. Developed by Google in
2014 and written in Go~\cite{k8sGit}, it allows different machines (called nodes
from now on) to create an abstraction layer and to form a cluster where is
possible to run Docker containers without bothering about hardware constraints.
Nodes can have specific roles and specialization, all of them managed by one or
many master nodes, that actually manage the orchestration. On top of that,
Kubernetes is virtualization-agnostic, meaning that it offers the possibility to
change the virtualization software in use (i.e. from Container-based technology
to Virtual machine software).
\begin{figure}[h]
 \centering \includegraphics[scale=0.35]{kubernetes_logo}
 \caption{Kubernetes Logo}
 \label{chap:intro:img:k8s_logo}
\end{figure}

@@

Another fundamental component in the architecture is the MANO, that has to
manage the Network Services and the deployment of new VNFs. Initially identified
with Openbaton (the component is described later
in~\ref{chap:prjan:sec:tech:sub:other:sub:openbaton}), afterword it has been
discovered to be not flexible enough to support other kinds of virtualization
techniques than pure or in-cloud virtual machines. Since this was a strong
project requisite that could not be easily changed, it has been decided to opt
for the creation of a smaller and simpler clone that offered a full integration
with Docker and Kubernetes.

\subsection{Available SFC solutions}
The tools analysis revealed little about already existent VNF or NS
implementations, so it was necessary to fill the gap. Since this one of our
thesis goals, we developed a complete solution to create a complete SFCs, trying
to standardize our vocabulary as much as possible, and giving basic
documentation when possible.

@@

\subsubsection{Docker-compose} 
Docker-compose is a tool that allows services orchestration. It automates most
of the tasks that should be performed by hand when launching one or multiple
docker containers. It's particularly useful when more containers have to
interact together, because all the services rules are described in a simple YAML
file. On top of that, it is possible to restart/stop/destroy selected services,
while keeping the other up. Additionally, it supports containers scaling and it
offers the possibility to simply get logs, interact of inspect with the defined
services. With the birth of Docker Swarm, work has started to port this tool to
the Swarm framework. At the time of this writing, the procedure to port a
Docker-compose configuration in Docker Swarm (which a description of the product
can be found in sub-section~\ref{chap:prjan:sec:tech:sub:other:sub:swarm}) it
has become completely transparent, bundling all the necessaries resources.

@@

\subsubsection{Openbaton}
\label{chap:prjan:sec:tech:sub:other:sub:openbaton}
Patronized by the Institute Fraunhofer, it is an open-source, customizable NFV
MANO-compliant framework, that offers many configurations options, and since
it's written in Java, there is the possibility to add components (plug-ins)
dynamically. Openbaton can be installed as a normal computer program (they offer
packages for the most common distros) but it can be deployed on Docker
containers too. It's designed to work with cloud providers like Amazon or Google
Compute engine, but it offers compatibility with on-premise cloud solutions like
Openstack, where it deploys virtual machines using the API the platform offers.
It can store VNF configurations saved as Topology and Orchestration
Specification for Cloud Applications (TOSCA) Yet Another Markup Language (YAML)
and it can handle multiple Point of Presence (PoP). It offers a VNF lifecycle
management out-of-the-box, with the possibility to customize it based on the
user needs. Its modular design allows developers to change parts of the codebase
with custom ones, making the product really flexible to different platforms.
Performing this operation requires a deep knowledge of the framework though.

>>>>

As described in chapter~\ref{chap:intro}, the creation of a full ETSI MANO
compliant architecture, following the suggestions described in RFC 7665,
requires the integration of multiple tools (some of them have been already
briefly introduced in~\ref{chap:intro}). The tools were chosen after their
analysis, where pro and cons have been studied and compared to the necessity of
this thesis. Since the project has been developed with the cloud provided by the
University of Padua, Openstack (which a product analysis can be found later
in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.

@@

\subsubsection{Openstack}
\label{chap:prjan:sec:openstack}
Created in 2010 by a joint operation between Rackspace Hosting and
NASA~\cite{openstackWebsite}, Openstack allows on-premise cloud installations,
using bare-metal resources to provide common cloud services as object storage,
virtual machine deploy and virtual networking. Furthermore its modularity offers
the possibility to add additional components, even proprietary, to achieve a
complete cloud solution.
\begin{figure}[h]
 \centering \includegraphics[scale=0.58]{openstack_logo}
 \caption{Openstack logo}
 \label{chap:prjan:img:openstack_logo}
\end{figure}

@@

As already introduced before, Kubernetes is an open source software framework
specialized in container management and orchestration. Developed by Google in
2014 and written in Go~\cite{k8sGit}, it allows different machines (called nodes
from now on) to create an abstraction layer and to form a cluster where is
possible to run Docker containers without bothering about hardware constraints.
Nodes can have specific roles and specializations, all of them managed by one or
many master nodes, that actually manage the orchestration. On top of that,
Kubernetes is virtualization-agnostic, meaning that it offers the possibility to
change the virtualization software in use (i.e. from Container-based technology
to Virtual machine one).
\begin{figure}[h]
 \centering \includegraphics[scale=0.35]{kubernetes_logo}
 \caption{Kubernetes Logo}
 \label{chap:intro:img:k8s_logo}
\end{figure}

@@

Another fundamental component in the architecture is the MANO, that has to
manage the Network Services and the deployment of new VNFs. Initially identified
with Openbaton (the component is described later
in~\ref{chap:prjan:sec:tech:sub:other:sub:openbaton}), afterword it has been
discovered to be not flexible enough to support other kinds of virtualization
techniques than pure or in-cloud virtual machines. Since this was a strong
project requisite that could not be easily changed, it has been decided to opt
for the creation of a smaller and simpler alternative that offered a full
integration with Docker and Kubernetes.

\subsection{Available SFC solutions}
The tools analysis revealed little about already existent VNF or NS reusable
framework implementations, so it was necessary to fill the gap. We then
developed a complete solution to create complete SFCs, trying to standardize
our vocabulary as much as possible, and giving basic documentation.

@@

\subsubsection{Docker-compose} 
Docker-compose is a tool that allows services orchestration. It automates most
of the tasks that should be performed by hand when launching one or multiple
docker containers. It is particularly useful when more containers have to
interact together, because all the services rules are described in a simple YAML
file. On top of that, it is possible to restart/stop/destroy selected services,
while keeping the other up. Additionally, it supports containers scaling and it
offers the possibility to simply get logs, interact of inspect with the defined
services. With the birth of Docker Swarm, work has started to port this tool to
the Swarm framework. At the time of this writing, the procedure to port a
Docker-compose configuration in Docker Swarm (which a description of the product
can be found in sub-section~\ref{chap:prjan:sec:tech:sub:other:sub:swarm}) it
has become completely transparent, bundling all the necessaries resources.

@@

\subsubsection{Openbaton}
\label{chap:prjan:sec:tech:sub:other:sub:openbaton}
Patronized by the Institute Fraunhofer, it is an open-source, customizable NFV
MANO-compliant framework, that offers many configurations options, and since
it's written in Java, there is the possibility to add components (plug-ins)
dynamically. Openbaton can be installed as a normal computer program (they offer
packages for the most common distros) but it can be deployed on Docker
containers too. It is designed to work with cloud providers like Amazon or
Google Compute engine, but it offers compatibility with on-premise cloud
solutions like Openstack, where it deploys virtual machines using the API the
platform offers. It can store VNF configurations saved as Topology and
Orchestration Specification for Cloud Applications (TOSCA) Yet Another Markup
Language (YAML) and it can handle multiple Point of Presence (PoP). It offers a
VNF lifecycle management out-of-the-box, with the possibility to customize it
based on the user needs. Its modular design allows developers to change parts of
the codebase with custom ones, making the product really flexible to different
platforms. Performing this operation requires a deep knowledge of the framework
though.
